MSDS6372_EDAProject2

### 2021-03-20 DataSelection_EDA.RMD
update DataSelection

2. Predicting if someone makes over 50k
https://archive.ics.uci.edu/ml/datasets/Adult
Pro: Easier to understand data, less variable, Train Test set already split and provided for you
Con: Have missing data - however due to the amount we can either delete or set them to unknown (should not affect result as much)
    Still have unbalance data for training dataset <=50k: 24720, >50kL 7841 - not as bad
                                  test dataset <=50K.:12435 >50K. : 3846
```{r}
setwd("/Users/mingyang/Desktop/SMU/Applied Statistics/MSDS6372Project2/Adult-Income")
library(tidyverse)
adult = read.csv("adult.data.csv",header=FALSE,sep=",",strip.white=TRUE) 
colnames(adult) <- c(
  "age", # continuous
  "workclass",# Private, Self-emp-not-inc, Self-emp-inc, Federal-gov, Local-gov, State-gov, Without-pay, Never-worked
  "fnlwgt", # continous, final weight. In other words, this is the number of people the census believes the entry represents
  "education", # Bachelors, Some-college, 11th, HS-grad, Prof-school, Assoc-acdm, Assoc-voc, 9th, 7th-8th, 12th, 
              # Masters, 1st-4th, 10th, Doctorate, 5th-6th, Preschool.education-num: continuous
  "education-num", # continuous
  "marital-status", # Married-civ-spouse, Divorced, Never-married, Separated, Widowed, Married-spouse-absent, Married-AF-spouse.
  "occupation",  # Tech-support, Craft-repair, Other-service, Sales, Exec-managerial, Prof-specialty, Handlers-cleaners, Machine-op-inspct,                 # Adm-clerical, Farming-fishing, Transport-moving, Priv-house-serv, Protective-serv, Armed-Forces
  "relationship", # Wife, Own-child, Husband, Not-in-family, Other-relative, Unmarried
  "race", # White, Asian-Pac-Islander, Amer-Indian-Eskimo, Other, Black
  "sex",   # Female, Male
  "capital-gain", # continuous
  "capital-loss", # continuous
  "hours-per-week", # continuous
  "native-country", # United-States, Cambodia, England, Puerto-Rico, Canada, Germany, Outlying-US(Guam-USVI-etc), India, Japan, Greece, South, China, Cuba, Iran, Honduras, Philippines, Italy, Poland, Jamaica, Vietnam, Mexico, Portugal, Ireland, France, Dominican-Republic, Laos, Ecuador, Taiwan, Haiti, Columbia, Hungary, Guatemala, Nicaragua, Scotland, Thailand, Yugoslavia, El-Salvador, Trinadad&Tobago, Peru, Hong, Holand-Netherlands
  "income" # expects to predict
)
str(adult)
## First, convert "?"s to NAs...
adult[adult == "?"] <- NA
variables.to.factor = c("workclass","education","marital-status","occupation","relationship","race","sex","native-country","income")
adult[variables.to.factor] = lapply(adult[variables.to.factor],factor)
summary(adult)

@@

adult.missing = adult %>% filter_all(any_vars(is.na(.)))
adult.test.missing = adult.test %>% filter_all(any_vars(is.na(.)))
view(adult.missing)
```

3. Assessing risk factors and predicting if a woman with osteopetrosis will have a bone fracture within the first year after joining the study.
Resource:R package aplore3, use the glow_bonemed data set
Pro: No missing data, already cleaned data for the most part, less many level categories
Con: Much smaller dataset - ratio of fracture is still about 3:1 No:Yes
```{r}
library(aplore3)
head(glow_bonemed)
str(glow_bonemed)
summary(glow_bonemed)
```

>>>>

2. Predicting if someone makes over 50k
https://archive.ics.uci.edu/ml/datasets/Adult
Pro: Easier to understand data, less variable, Train Test set already split and provided for you
Con: Have missing data - however due to the amount we can either delete or set them to unknown (should not affect result as much)
    Still have unbalance data for training dataset <=50k: 24720, >50kL 7841 - not as bad
                                  test dataset <=50K.:12435 >50K. : 3846
```{r}
setwd("/Users/mingyang/Desktop/SMU/Applied Statistics/MSDS6372Project2/Adult-Income")
library(tidyverse)
adult = read.csv("adult.data.csv",header=FALSE,sep=",",strip.white=TRUE) 
colnames(adult) <- c(
  "age", # continuous
  "workclass",# Private, Self-emp-not-inc, Self-emp-inc, Federal-gov, Local-gov, State-gov, Without-pay, Never-worked
  "fnlwgt", # continous, final weight. In other words, this is the number of people the census believes the entry represents
  "education", # Bachelors, Some-college, //11th, HS-grad, Prof-school, Assoc-acdm, Assoc-voc, 9th, 7th-8th, 12th, 
              # Masters, 1st-4th, 10th, Doctorate, 5th-6th, Preschool.education-num: continuous
  "education-num", # continuous
  "marital-status", # Married-civ-spouse, Divorced, Never-married, Separated, Widowed, Married-spouse-absent, Married-AF-spouse.
  "occupation",  # Tech-support, Craft-repair, Other-service, Sales, Exec-managerial, Prof-specialty, Handlers-cleaners, Machine-op-inspct,                 # Adm-clerical, Farming-fishing, Transport-moving, Priv-house-serv, Protective-serv, Armed-Forces
  "relationship", # Wife, Own-child, Husband, Not-in-family, Other-relative, Unmarried
  "race", # White, Asian-Pac-Islander, Amer-Indian-Eskimo, Other, Black
  "sex",   # Female, Male
  "capital-gain", # continuous
  "capital-loss", # continuous
  "hours-per-week", # continuous
  "native-country", # United-States, Cambodia, England, Puerto-Rico, Canada, Germany, Outlying-US(Guam-USVI-etc), India, Japan, Greece, South, China, Cuba, Iran, Honduras, Philippines, Italy, Poland, Jamaica, Vietnam, Mexico, Portugal, Ireland, France, Dominican-Republic, Laos, Ecuador, Taiwan, Haiti, Columbia, Hungary, Guatemala, Nicaragua, Scotland, Thailand, Yugoslavia, El-Salvador, Trinadad&Tobago, Peru, Hong, Holand-Netherlands
  "income" # expects to predict
)
str(adult)
## First, convert "?"s to NAs...
adult[adult == "?"] <- NA
variables.to.factor = c("workclass","education","marital-status","occupation","relationship","race","sex","native-country","income")
adult[variables.to.factor] = lapply(adult[variables.to.factor],factor)
summary(adult)

@@

adult.missing = adult %>% filter_all(any_vars(is.na(.)))
adult.test.missing = adult.test %>% filter_all(any_vars(is.na(.)))
# Look at missing data
library(naniar)
gg_miss_var(adult)
gg_miss_var(adult.test)
```

3. Assessing risk factors and predicting if a woman with osteopetrosis will have a bone fracture within the first year after joining the study.
Resource:R package aplore3, use the glow_bonemed data set
Pro: No missing data, already cleaned data for the most part, less many level categories
Con: Much smaller dataset - ratio of fracture is still about 3:1 No:Yes
```{r}
library(aplore3)
head(glow_bonemed)
str(glow_bonemed)
summary(glow_bonemed)
gg_miss_var(glow_bonemed)
```
